### 2019-05-20 Datelife_paper1.Rmd
figure cap format

---
output: pdf_document
fontsize: 12pts
textalign: left  # this does not work...
geometry: margin = 1in
bibliography: library_red.bib
csl: systematic-biology.csl
link-citations: yes
# indent: true # this works
# header-includes:
#    - \setlength\parindent{24pt}  # this works too, but I'm gonna indent manually with 4 spaces
# nope, 4 spaces indents are not recognized, they do something else
linestretch: 2
header-includes:
    - \usepackage[left]{lineno}
    - \linenumbers
---
```{r eval = TRUE, echo = FALSE}
source("Datelife_paper1_global.R")
```

@@

<!--(BOLD and OToL species names, are they homogeneous/standardized?-->
The basic `DateLife` workflow is shown in `r fig_workflow_main` and consists of:

@@

Source chronogram summary tree can be assembled using the Super Distance Matrix
(SDM) supertree construction approach [@Criscuolo2006] or using the median of branch
lengths and the hierarchical clustering method<!-- of ttol-->. <!--Also, I discovered the median gives non ultrametric trees... but I thonk it is fixed now, check it!).-->
Tree dating and simulation options are performed with various algorithms:
Branch Length Adjuster (BLADJ) is a simple algorithm to distribute ages of undated
nodes evenly, which minimizes age variance in the chronogram [@Webb2008].
PATHd8 is a non-clock, rate-smoothing method [@Britton2007] to date trees.
treePL, is a semi-parametric, rate-smoothing, penalized likelihood dating method
[@Smith2012].
MrBayes [@Huelsenbeck2001; @Ronquist2003] can be used when adding taxa at
random, following a reference taxonomy or a topological constraint. It draws ages
from a pure birth model, as implemented by Jetz and collaborators [-@Jetz2012].
To apply calibrations to a tree, th econgruification algorithm described in [@Eastman2013]
is used to find shared nodes between trees (congruent nodes).

@@

`DateLife`'s code speed was tested on an Apple iMac
with one 3.4 GHz Intel Core i5 processor. <!--No multiprocessor tests were performed-->
We registered variation in computing time relative to number of input names
<!--number of chronograms in database--> and `DateLife` service.
Input processing increases roughly linearly with number of input taxon names, and
increases considerably if tnrs service is activated (`r fig_runtime_main`).
<!--
Speed with different number of input lineages and types of analysis:
-->
Results show that searching time increases linearly with number of input names and
number of chronograms in database.

@@

<!--
Speed of web interface and of r package (in computers with different capacities?)
-->
`DateLife`'s code performance was evaluated with a set of unit tests designed and
implemented with the R package testthat [@RCoreTeam2018]. These tests were run both
locally –using the devtools package [@RCoreTeam2018]– and on a public server –via
GitHub– using the continuous integration tool Travis CI (<//travis-ci.org>). At
present, unit tests cover around 30% (for now) of `DateLife`'s code (<https://codecov.io/gh/phylotastic/datelife>). <!-- unit test make code robust, verifiable, debuggable, -->

@@

Find a clade with at least one chronogram containing all clade's species.
(Penguins look good, but they are giving weird results in SDM)

@@

Use ltts to compare for now.
Fig. X2 shows comparison of available chronograms for Felidae species and chronograms
generated through DateLife

think of a test to compare trees, topology- and date-wise

<!--Bird (or reptile) chronograms, too long time...
finches is good
nothofagus
Look for all chronograms containing any birds
Or, look for chronograms containing small and old lineages
Determine which clade of birds has the more chronograms (have been dated
more times) and use that as biological example-->

@@

<!--**Acknowledgements**-->
<span style="font-variant:small-caps;">
Acknowledgements
</span>
We thank colleagues (students and postdocs) at the O'Meara Lab at the University
of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used for DateLife.
The many scientists who publish their chronograms in an open, reusable form, and
the scientists who curate them for deposition in OpenTree.
The US National Science Foundation (NSF) for funding nearly all the above, in addition
to the ABI grant that funded this project itself.

<!--**References**-->
<span style="font-variant:small-caps;">
References
</span>

>>>>

---
output: 
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
    includes:
      # in_header: preamble.tex
      after_body: [Datelife_paper1_figures_noyaml.tex]
    # fig_caption: no
textalign: left  # this does not work...
geometry: margin = 1in
bibliography: library_red.bib
csl: systematic-biology.csl
link-citations: yes
# indent: true # this works
# header-includes:
#    - \setlength\parindent{24pt}  # this works too, but I'm gonna indent manually with 4 spaces
# nope, 4 spaces indents are not recognized, they do something else
linestretch: 2
header-includes:
    # - \usepackage{caption}
    - \usepackage[left]{lineno}
    - \linenumbers  
  # - \documentclass[oupdraft]{sysbio} # this is duplicated in tex generated by knitr
  # - \usepackage{xr}
  # - \externaldocument{Datelife_paper1_figures}
    # - \usepackage[utf8]{inputenc}
    # - \usepackage[english]{babel}
    # - \usepackage[document]{ragged2e}
---

```{r eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE}
source("Datelife_paper1_global.R")
```

@@

<!--(BOLD and OToL species names, are they homogeneous/standardized?-->
The basic `DateLife` workflow is shown in figure \ref{fig:workflow} and consists of:

@@

Source chronogram summary tree can be assembled using the Super Distance Matrix
(SDM) supertree construction approach [@Criscuolo2006] or using the median of branch
lengths and the hierarchical clustering method<!-- of ttol-->. <!--Also, I discovered the median gives non ultrametric trees... but I think it is fixed now, check it!).-->
Tree dating and simulation options are performed with various algorithms:
Branch Length Adjuster (BLADJ) is a simple algorithm to distribute ages of undated
nodes evenly, which minimizes age variance in the chronogram [@Webb2008].
PATHd8 is a non-clock, rate-smoothing method [@Britton2007] to date trees.
treePL, is a semi-parametric, rate-smoothing, penalized likelihood dating method
[@Smith2012].
MrBayes [@Huelsenbeck2001; @Ronquist2003] can be used when adding taxa at
random, following a reference taxonomy or a topological constraint. It draws ages
from a pure birth model, as implemented by Jetz and collaborators [-@Jetz2012].
To apply calibrations to a tree, th econgruification algorithm described in [@Eastman2013]
is used to find shared nodes between trees (congruent nodes).

@@

`DateLife`'s code speed was tested on an Apple iMac
with one 3.4 GHz Intel Core i5 processor. <!--No multiprocessor tests were performed-->
We registered variation in computing time relative to number of input names
<!--number of chronograms in database--> and `DateLife` service.
Input processing increases roughly linearly with number of input taxon names, and
increases considerably if tnrs service is activated (Fig. \ref{fig:runtime1}).
<!--
Speed with different number of input lineages and types of analysis:
-->
Results show that searching time increases linearly with number of input names and
number of chronograms in database.

@@

<!--
Speed of web interface and of r package (in computers with different capacities?)
-->
`DateLife`'s code performance was evaluated with a set of unit tests designed and
implemented with the R package testthat [@RCoreTeam2018]. These tests were run both
locally –using the devtools package [@RCoreTeam2018]– and on a public server –via
GitHub– using the continuous integration tool Travis CI (<//travis-ci.org>). At
present, unit tests cover more than 50% (for now) of `DateLife`'s code (<https://codecov.io/gh/phylotastic/datelife>). <!-- unit test make code robust, verifiable, debuggable, -->

@@

<!--Bird (or reptile) chronograms, too long time...
finches is good
nothofagus
Look for all chronograms containing any birds
Or, look for chronograms containing small and old lineages
Determine which clade of birds has the more chronograms (have been dated
more times) and use that as biological example. Fringillidae is a good candidate
Find a clade with at least one chronogram containing all clade's species.
(Penguins look good, but they are giving weird results in SDM)

@@

Use ltts to compare for now.

think of a test to compare trees, topology- and date-wise
-->
In this section we demonstrate the types of outputs that can be obtained with `datelife`, using the bird family Fringillidae of true finches as example. We performed a higher-taxon search to obtain all data on lineage divergence available from `datelife`'s database for all recognised species within the Fringillidae (`r length(dq$cleaned_names)` spp. according to the Open Tree of Life taxonomy). We found `r length(dr)` trees across `r length(unique(names(dr)))` studies (Fig. \ref{fig:schronograms}.

<!--Show code here, in the paper? Or show it only in the examples? Or only in the vignette? Write down what about datelife will be shown in each of these.
Paper should show outputs from web site and package. 
Vignette will show the R code used to generate the outputs displayed here and in the website.
datelife examples will explain more on what is done in every step.
-->

@@

<!--**Acknowledgements**-->
<span style="font-variant:small-caps;">
Acknowledgements
</span>

We thank colleagues (students and postdocs) at the O'Meara Lab at the University
of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used for DateLife.
The many scientists who publish their chronograms in an open, reusable form, and
the scientists who curate them for deposition in OpenTree.
The US National Science Foundation (NSF) for funding nearly all the above, in addition
to the ABI grant that funded this project itself.

\begin{center}
\textsc{References}
\end{center}

### 2019-05-20 Datelife_paper1_figures.Rmd
figure cap format

---
output: pdf_document
fontsize: 12pts
textalign: left  
geometry: margin = 1in
bibliography: library_red.bib
csl: systematic-biology.csl
link-citations: yes
linestretch: 2
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
---

```{r setup, echo = FALSE}
source("Datelife_paper1_global.R")
```

```{r pressure, out.width="100%", include=TRUE, fig.align="center", fig.cap=fig_workflow_main_foot, echo=FALSE}
knitr::include_graphics("fig1.pdf", dpi = 560)
```

```{r eval = TRUE, results = 'asis', out.width="100%", include=TRUE, fig.align="center", fig.cap=fig_runtime_main_foot, echo=FALSE}
download.file(url = "https://github.com/phylotastic/datelife/raw/master/data-raw/benchmark/runtime_tests/2_tests/2_random_spp_names/mdq_plus_gdr_2018.05.29_fill.pdf", destfile = "fig_runtime_main.pdf", mode="wb")
knitr::include_graphics("fig_runtime_main.pdf", dpi = 560)

```

>>>>

---
output: 
  bookdown::pdf_document2:
    keep_tex: yes
textalign: left
geometry: margin = 1in
bibliography: library_red.bib
csl: systematic-biology.csl
link-citations: yes
linestretch: 2
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{caption}
  # - \usepackage{xr}
  #- \externaldocument{name of external document to reference fugures from there in here}
---

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
source("Datelife_paper1_global.R")
```

<!-- ```{r myfigure, fig.cap = "My figure"} -->
<!-- # fig.show = "hide" AND include = FALSE # does not generate the label for the figure -->
<!-- # fig.cap = " " inactivates the label reference -->
<!-- plot(1:10) -->
<!-- ``` -->

<!-- My plot is called \@ref(fig:myfigure). -->

<!-- \begin{figure}[!ht] -->
<!-- \includegraphics{blank.png} -->
<!-- \caption{} -->
<!-- \label{fig:workflow} -->
<!-- \end{figure} -->
<!-- Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with DateLife, via the R package or through the website. Details on the functions involved on each workflow are shown in `datelife`'s R package vignette. -->

<!-- \begin{figure}[!ht] -->
<!-- \includegraphics{blank.png} -->
<!-- \caption{} -->
<!-- \label{fig:runtime1} -->
<!-- \end{figure} -->
<!-- Computation time of input processing and search across `datelife`s  -->
<!--   chronogram database. -->

<!-- \begin{figure}[!ht] -->
<!-- \includegraphics{blank.png} -->
<!-- \caption{} -->
<!-- \label{fig:schronograms} -->
<!-- \end{figure} -->
<!-- Lineage through time (LTT) plots of Fringillidae (bird family of true finches) source chronograms. The latter are obtained by searching a chronogram database and subsetting chronograms conaning at least two species. Arrows indicate maximum age of each source chronogram and numbers indicate the reference of original studies 1: @barker2012going, 2: @barker2015new, 3: @burns2014phylogenetics, 4: @claramunt2015new, 5: @gibb2015new, 6: @Hedges2015, 7: @hooper2017chromosomal, 8: @Jetz2012, 9: @price2014niche. -->

<!-- \begin{figure}[!ht] -->
<!-- \includegraphics{blank.png} -->
<!-- \caption{} -->
<!-- \label{fig:cvbladj} -->
<!-- \end{figure} -->
<!-- LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarizing information from source chronograms found for the Fringillidae. Arrows indicate maximum age. -->

<!-- \newpage -->

Figure 1. 
Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with DateLife, via the R package or through the website. Details on the functions involved on each workflow are shown in `datelife`'s R package vignette.

Figure 2. 
Computation time of input processing and search across `datelife`s chronogram database.

Figure 3. 
Lineage through time (LTT) plots of source chronograms containing all or a subset of species from the bird family Fringillidae of true finches. Arrows indicate maximum age of each chronogram. Numbers reference to chronograms' original publications 1: @barker2012going, 2: @barker2015new, 3: @burns2014phylogenetics, 4: @claramunt2015new, 5: @gibb2015new, 6: @Hedges2015, 7: @hooper2017chromosomal, 8: @Jetz2012, 9: @price2014niche.

Figure 4.
LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarizing information from source chronograms found for the Fringillidae. Arrows indicate maximum age.

\newpage

\begin{figure}[!h]
\includegraphics{Fig1.pdf}
\caption{}
\label{fig:workflow}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics{fig_runtime1.pdf}
\caption{}
\label{fig:runtime1}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics{fig_schronograms1.pdf}
\caption{}
\label{fig:schronograms}
\end{figure}

\newpage

\begin{figure}[!ht]
\includegraphics{fig_crossval_bladj.pdf}
\caption{}
\label{fig:cvbladj}
\end{figure}
